Utilizing Parallel Workers: \\LLVM's Vectorization Plan

### 2024-11-22 paper/paper.tex
paper

\section{LLVM's Vectorization Plan}
\label{sec:vplan}
Originally LLVM implements two different vectorizers for auto-vectorization, LoopVectorize (LV)
and SLPVectorize. These vectorizers are a part of LLVM's middle-end and run after module 
simplification. As VPlan has not yet been integrated into SLP-Vectorization we will only focus on
explaining the original LoopVectorize structure.

LoopVectorize was designed to focus only on vectorization of innermost loops~\cite{llvmintrvplan}. 
When starting with a initial scalable loop in IR it will first run a legality phase to check 
whetever and how it is legal to vectorize the loop. This phase would produce artifacts/maps like if 
and what runtime aliasing checks would be required. If some vectorization would be legal the 
cost model would run next, creating more artifacts based on the scalar IR and legality artifacts. 
These cost model artifacts would associate different vectorization factors with a cost.
Lastly the transformation phase would vectorize the underlying IR based on all the produced 
artifacts.

However, this rigid system of LoopVectorize was a problem. Extending LV to include outer loop 
and nested loop vectorizations would have been very challenging. Producing and considering more and 
more artifacts when trying to transform the scalar loop into a vectorized one all at once did not 
scale well.

@@

\subsubsection{Function Vectorization}
As far as we now there is currently no active development for function vectorization in the
llvm-project's upstream yet.

>>>>

\section{LLVM's Vectorization Plan}
\label{sec:vplan}

\subsection{History}
LLVM implements two different vectorizers for auto-vectorization, LoopVectorize (LV)
and SLPVectorize. These vectorizers are a part of LLVM's middle-end and run after module 
simplification has taken place. As VPlan has not yet been integrated into SLP-Vectorization 
we will only focus on explaining the original LoopVectorize structure.

LoopVectorize was designed to focus only on vectorization of innermost loops~\cite{llvmintrvplan}. 
When starting with a initial scalable loop in IR it would first run a legality phase to check 
whetever and how it is legal to vectorize the loop. This phase would produce artifacts/maps like if 
and what runtime aliasing checks would be required. Then the cost model ran next, creating more 
artifacts based on the scalar IR and legality artifacts. These cost model artifacts would 
associate different vectorization factors with a cost. Lastly the transformation phase would 
vectorize the underlying IR in one go based on all the produced artifacts.

However, this rigid flow of LoopVectorize was a problem. Extending LV to include outer loop 
and nested loop vectorizations would have been very challenging, as LV was designed to flatten a
loops control flow into a single basic block. Producing and considering more and 
more artifacts when trying to transform the scalar loop into a vectorized one all at once did also 
not scale well.

@@

\subsubsection{Function Vectorization}
As far as we now there is currently no active development for whole-function vectorization in the
llvm-project's upstream yet.
